---
title: 'Project B1'
output: github_document
---

# Exercise 1 and 2: Making and Documenting a Function

We'll write a function `na_summary()` and document it using roxygen2 tags.

```{r}
#' Missing and Non-Missing Values Summary
#' Summarizes the count and percentage of missing and non-missing values in each column of a data frame.
#' 
#' @param df An object of class "data.frame". Named df as short for data frame.
#' @return An object of class "data.frame" in a long data format that summarizes the count and percentage of missing and non-missing values in each column of a data frame.
#' @export 

na_summary <- function(df){
  if(!require(tidyverse)){
    stop("The tidyverse packages must be installed. Run install.packages(\"tidyverse\") and then try again.")
  }
  if(!is.data.frame(df)){
    stop("This function only works for object of class \"data.frame\" \n", "You have provided an object of class:", class(df))
  }
  if(nrow(df) == 0 | ncol(df) == 0){
    stop("This function only works for \"data.frame\" object of non-zero dimensions")
  }
  df %>%
    gather(key = "key", value = "val") %>%
    mutate(is_na = is.na(val)) %>%
    group_by(key) %>%
    mutate(total = n()) %>%
    group_by(key, total, is_na) %>%
    summarise(na_count = n()) %>%
    mutate(na_pct = na_count/total *100)
}
```


# Exercise 3: Demonstrating Examples

The function `na_summary()` was inspired by the tasks of exploring many datasets in the `datateachr` package. 
We'll run `na_summary()` on the `apt_buildings` and `vancouver_trees` datasets from `datateachr`. 

```{r error = TRUE}
# Load packages  ---------------------------
library(datateachr)

# Run na_summary() on data ---------------------------
# Complete dataset
na_summary(apt_buildings)
na_summary(vancouver_trees)

# Subset of the dataset
apt_buildings %>%
    select(ward, prop_management_company_name, year_built, no_of_storeys, no_of_units, property_type, no_barrier_free_accessible_units, no_of_accessible_parking_spaces, heating_type) %>%
  na_summary()


```
Notice that `na_summary()` works on a data frame with columns of various object classes. 

```{r}
glimpse(apt_buildings)
glimpse(vancouver_trees)
```

We can also create a new data frame or one from an existing object to run `na_summary()`.

```{r}
# Create a non data frame object ---------------------------
# matrix array object
alpha_num <- cbind(col1= 1:3, col2 = c("A", "B", "C"))
alpha_num
class(alpha_num)

# Convert into data frame object ---------------------------
alpha_num_df <- data.frame(alpha_num)

# Run na_summary() ---------------------------
na_summary(alpha_num_df)
```

Overall, `na_summary()` is helpful for exploring data and making informed decisions on if/how we should use variables. 





# Exercise 4: Testing Function

We'll use the `testthat` package to run formal tests for our function. 

```{r}
# Load packages  ---------------------------
library(testthat)
```

**Test 1**
We expect an error message when `na_summary()` is used on an object that is not of class data.frame. 
We evaluated this on objects of class character, integer, and matrix array. 
```{r}
test_that("Function returns error if df is not a data.frame object", {
  expect_error(na_summary(letters)) 
  expect_error(na_summary(1:26))
  expect_error(na_summary(cbind(letters, 1:26)))
})
```

**Test 2**
We expect an error message when `na_summary()` is used on a data.frame object of zero dimension. 
We evaluated this on a dataframe with zero dimension only of the row, only of the column, and of both the column and row. 

```{r}
test_that("Function returns error when df has zero dimension", {
  expect_error(na_summary(data.frame()))
  expect_error(na_summary(data.frame(col1 = numeric(0), col2 = character(0))))
  expect_error(na_summary(data.frame(numeric(0))))
})
```

**Test 3**
We expect `na_summary()` to return a data.frame of summarized missing and non-missing values on a data.frame with columns of mixed class types; numeric and character. 

```{r}
test_that("Function returns dataframe of missing and non-missing values summary when df consists of numeric and character class types", {
  expect_s3_class(na_summary(data.frame(col1=("C, B, A"), col2=c("A", "NA", "C"))), "data.frame")
  expect_s3_class(na_summary(data.frame(col1=1:3, col2= c(4, NA, 5))), "data.frame")
})

```

**Test 4**
We expect `na_summary()` to return a data.frame of summarized missing and non-missing values on a data.frame consisting of entirely all NA or none NA values. 

```{r}
test_that("Function returns data frame of missing and non-missing values summary when df consists entirely of NAs or no NAs", {
  expect_s3_class(na_summary(data.frame(col1=1:3, col2=c("A", "B", "C"))), "data.frame")
  expect_s3_class(na_summary(data.frame(col1=1:3, col2=c(NA, NA, NA))), "data.frame")
})
```

