---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nasummary

<!-- badges: start -->
<!-- badges: end -->

The goal of `nasummary` is to summarize the count and percentage of missing and non-missing values for each columns of a data frame. When working with new data, it is good practice to perform an ‘exploratory data analysis’ (EDA) which refers to the initial exploration of the data to understand patterns, anomalies, and check assumptions. Often in the EDA, the number of missing values in a variable column followed by a calculation of the percentage of data that the missing values account for is explored. This is to determine whether a variable should be used in further analyses. For example, based on your knowledge of the data and modelling assumptions, you have decided to include variables that contain missing values only if it accounts for less than 25% of the total data.  The `nasummary` package is designed to simplify this calculation by returning a data frame that summarizes the count and percentage of missing and non-missing values in an easy to read long data format.  

## Installation

You can install the development version of nasummary from [GitHub](https://github.com/): 

```{r message=FALSE}
# install.packages("devtools")
devtools::install_github("stat545ubc-2022/assignment-b1-and-b2-vinky-wang/nasummary/")
```
**Note**: uncomment install.packages("devtools") if you do not already have `devtools` installed. 



## Example

We can create a data frame then run `na_summary()` to obtain the count and percentage of missing and non-missing values. 

```{r}
# Create a non data frame object ---------------------------
alpha_num <- cbind(col1= 1:3, col2 = c("A", "B", "C"))

# Convert into data frame object ---------------------------
alpha_num_df <- data.frame(alpha_num)

# Run na_summary() ---------------------------
library(nasummary)
na_summary(alpha_num_df)
```

